Child care situation of working mothers in Ghana 1998

### 2022-04-03 outputs/paper.Rmd

---
title: "Child care situation of working mothers in Ghana 1998"
author: 
  - Pascal Lee Slew
  - Yunkyung Park
thanks: "Code and data are available at: https://github.com/Pascal-304/dhs_analysis."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Working mothers are often faced with the conflict of responsibility towards their children and the role they play at work. The Ghana Demographic and Health Survey (GDHS) forms a part of a global initiative to collect data on family, fertility, and child health. In this paper, we used the 1998 GDHS and focused on the child care and working mothers aspect. Geographics and other aspects such as education are used to predict child care preference. We found that families living in Greater Accra are more likely to send their children to school and the majority of working mothers look after their children by themselves. The analysis of the child care and working mothers data provides insight into current trends and has significant implications for public policy and program decisions."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

This paper uses the Childcare While Working Data [@ghana1998] obtained from the 1998 Ghana Demographic and Health Survey (GDHS). The dataset was collected by downloading the final report of the 1998 GDHS published in the Demographic and Health Surveys (DHS) programme. The package 'pdftools' [@citepdftools] is used to download and read the content into R by parsing the pdf. The package 'stringi' [@citestringi] and 'tidyverse' [@citetidyverse] are then used to gather the data.

@@

The questionnaires were made available in the five major local languages (Akan, Ga, Ewe, Hausa, and Dagbani), making the survey accessible to a larger population. After the household listing, sports checks were conducted to make sure the work has been done correctly. Before the actual survey, a test was conducted and adjustments were made to the questionnaires based on the feedback obtained from the test. The questions were comprehensive and detailed to collect information on the socio-economic status of the household.

@@

Since the questionnaires were only available in the five major local languages, if the selected household was from a minor ethnic group and know only their own language, then it would be difficult for the interviewee to understand and answer the survey. Visitors who stayed a night before in the selected household were eligible for the interview. The visitor may be from a different region, so the targeted sample size for each region may be skewed.

@@

```{r generalchildcare, echo=FALSE, fig.cap="Distribution of the caretaker of the child while mother is working", fig.width=6, fig.height=4}
totl <-
  ghana %>%
  filter(character=="Total") %>%
  mutate(
    no_child_under_six = round(no_child_under_six/100 * number_of_employed_women, 0),
    one_more_child_under_six = round(one_more_child_under_six/100 * number_of_employed_women, 0)
  )
# make dataset in long format
totl_long <-
  totl %>%
  select(-c(character, total, caretaker_missing)) %>%
  pivot_longer(!background, names_to = "variable", values_to = "count")

@@

# plot the bar graph
totl_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") +
  geom_text(aes(label=paste(percentage, '%'), y=percentage+2),
            size = 3,
            position = position_dodge(width=0.5)) +
  theme_minimal() +
  labs(x='Percentage', y='Caretaker while at work') +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other")
  )
```

@@

```{r map, message=FALSE, echo = FALSE, warning=FALSE, fig.cap="Ghana administrative regions"}
# install.packages('rnaturalearth')
# install.packages('sf')
library(rnaturalearth)
library(sf)
# devtools::install_github("ropensci/rnaturalearthhires")
# The st_centroid function takes the sf object and returns the center of 
# the polygon, again as a sf object.
map_ghana <- ne_states(country = 'Ghana', 
                       returnclass = 'sf')  # grab provinces within Ghana

@@

# plot bar graph
resid_long %>% 
  filter(variables == 'No child under six' | variables == 'One or more child under six') %>% 
  ggplot(aes(x = background, 
             y = percentage,
             fill = variables)) +
  geom_bar(stat = 'identity',
           position = 'dodge') +
  labs(x = "Area",
       y = "Percentage",
       fill = "Child under six") +
  theme_minimal() +
  scale_fill_brewer() +
  theme(legend.position = "bottom")
```

@@

# plot bar graph
caretaker %>% 
  ggplot(aes(x = caretakers, 
             y = percentage,
             fill = background)) +
  geom_bar(stat = 'identity',
           position = "dodge") +
  labs(x = "Caretaker while at work",
       y = "Percentage",
       fill = "Area") +
  theme_minimal() +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

@@

# plot bar graph for caretaker comparisons and administrative regions
child %>% 
  filter(variable != 'number of employed women', variable != 'no child under six', variable != 'one or more child under six') %>% 
  ggplot(aes(x = variable,
             y = percentage,
             fill = background)) +
  geom_bar(stat = 'identity') + 
  labs(x = "Caretaker while at work",
       y = "Percentage") +
  theme_minimal() +
  coord_flip() +
  facet_wrap(~background) +
  theme(legend.position = "none")
```

@@

```{r chstatus, echo = FALSE, fig.cap = "Comparing the child status in different administrative regions"}
# looking at the number of employed women selected in the different administrative regions
# with the child under six
child %>%
  filter(variable=='no child under six'|variable=='one or more child under six') %>% 
  ggplot(aes(x = background,
             y = percentage,
             fill = variable)) +
  geom_bar(stat = 'identity', position = "dodge", width = 0.8) +
  labs(x = "Administrative regions",
       y = "Percentage",
       fill = "Child under six") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  coord_flip() 
```

@@

```{r motheredu, echo=FALSE, fig.cap="Mother's Education level", fig.width=6, fig.height=4}
# select mother edu
mother_edu <-
  ghana %>%
  filter(character=="Mother\'s_education") %>%
  mutate(
    no_child_under_six = round(no_child_under_six/100 * number_of_employed_women, 0),
    one_more_child_under_six = round(one_more_child_under_six/100 * number_of_employed_women, 0)
  )

@@

edu_long %>%
  filter(variable == 'no_child_under_six' | variable == 'one_more_child_under_six') %>%
  ggplot(aes(x = background, y = count, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Mother's education", y = "Number of employed women", fill = "Child under six") +
  theme_minimal() +
  scale_fill_discrete(labels = c("No child under six", 
                                 "One or more child under six")) +
  scale_x_discrete(labels = c("No_education" = "No education")) +
  theme(legend.position = 'bottom')
```

@@

# plot the bar graph
edu_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") + 
  theme_minimal() +
  labs(x='Caretaker while at work', y="Percentage") +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired help",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other")
  ) +
  facet_wrap(vars(background),
             dir = "v",
             scales = "free",
             labeller = labeller(background = 
                                   c("No_education" = "No education",
                                     "Primary" = "Primary",
                                     "Middle" = "Middle",
                                     "Secondary+" = "Secondary+")
                                 )
  )
```

@@

```{r workstatus, echo=FALSE, message=FALSE, fig.cap="Work status", fig.width=6, fig.height=4}
# select work status
work_status <-
  ghana %>%
  filter(character=="Work_status") %>%
  mutate(
    no_child_under_six = round(no_child_under_six/100 * number_of_employed_women, 0),
    one_more_child_under_six = round(one_more_child_under_six/100 * number_of_employed_women, 0)
  )

@@

work_long %>%
  filter(variable == 'no_child_under_six' | variable == 'one_more_child_under_six') %>%
  ggplot(aes(x = background, y = count, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Work status", y = "Number of employed women", fill = "Child under six") +
  theme_minimal() +
  scale_x_discrete(labels = c("For_family_member" = "For family member",
                              "For_someone_else" = "For someone else")) + 
  scale_fill_discrete(labels = c("No child under six", 
                                 "One or more child under six")) +
  theme(legend.position='bottom')
```

@@

# plot the bar graph
work_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") + 
  theme_minimal() +
  labs(x='Caretaker while at work', y="Percentage") +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired help",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other")
  ) +
  facet_wrap_paginate(vars(background),
                      ncol=2,
                      nrow=2,
                      dir = "h",
                      scales = "free",
                      labeller = labeller(background = c("For_family_member" = "For family member",
                                                         "For_someone_else" = "For someone else",
                                                         "Self-employed" = "Self-employed")
                      )
  )

@@

```{r occupation, echo=FALSE, message=FALSE, fig.cap="Occupation", fig.width=6, fig.height=4}
# select occupation
occupation <-
  ghana %>%
  filter(character=="Occupation") %>%
  mutate(
    no_child_under_six = round(no_child_under_six/100 * number_of_employed_women, 0),
    one_more_child_under_six = round(one_more_child_under_six/100 * number_of_employed_women, 0)
  )

@@

occup_long %>%
  filter(variable == 'no_child_under_six' | variable == 'one_more_child_under_six') %>%
  ggplot(aes(x = background, 
             y = count,
             fill = variable)) +
  geom_bar(stat = "identity",
           position = "dodge") +
  labs(x = "Sector",
       y = "Number of employed women",
       fill = "Child under six") +
  theme_minimal() +
  scale_fill_discrete(labels = c("No child under six", 
                                 "One or more child under six")) +
  theme(legend.position = 'bottom')
```

@@

# plot bar graph
occup_caretaker %>%
  ggplot(aes(x=caretaker, y=percentage, fill=background)) +
  geom_bar(stat = "identity", position="dodge") + 
  labs(x='Caretaker while at work', y="Percentage", fill="Occupation") +
  theme_minimal() +
  theme(legend.position='bottom') +
  coord_flip() +
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent", 
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired help",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other")
  )
```

@@

```{r employmentstatus, echo=FALSE, message=FALSE, fig.cap="Employment status", fig.width=6, fig.height=4}
# select employment status
employment_status <-
  ghana %>%
  filter(character=="Employment_status") %>%
  mutate(
    no_child_under_six = round(no_child_under_six/100 * number_of_employed_women, 0),
    one_more_child_under_six = round(one_more_child_under_six/100 * number_of_employed_women, 0)
  )

@@

emply_long %>%
  filter(variable == 'no_child_under_six' | variable == 'one_more_child_under_six') %>%
  ggplot(aes(x = background, y = count, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Employment status", y = "Number of employed women", fill = "Child under six") +
  theme_minimal() +
  scale_fill_discrete(labels = c("No child under six", 
                                 "One or more child under six")) +
  scale_x_discrete(labels = c("All_year_full_week" = "All year, full week",
                              "All_year_part_week" = "All year, part week")) +
  theme(axis.text.x = element_text(angle=15), legend.position = 'bottom')
```

@@

# plot the bar graph
emply_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") + 
  theme_minimal() +
  labs(x='Caretaker while at work', y="Percentage") +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired help",
        "child_in_school" = "schooling",
        "other_child" = "siblings",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other")
  ) +
  facet_wrap(vars(background),
             dir = "v",
             scales = "free",
             labeller = labeller(background = c("All_year_full_week" = "All year, full week",
                                                "All_year_part_week" = "All year, part week",
                                                "Seasonal" = "Seasonal",
                                                "Occasional" = "Occasional")
                                 )
  )

>>>>

---
title: "Child care situation of working mothers in Ghana 1998"
author: 
  - Pascal Lee Slew
  - Yunkyung Park
thanks: "Code and data are available at: https://github.com/Pascal-304/dhs_analysis."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Working mothers are often faced with the conflict of responsibility towards their children and the role they play at work. The Ghana Demographic and Health Survey (GDHS) forms part of a global initiative to collect data on family, fertility, and child health. In this paper, we used the 1998 GDHS and focused on the child care and working mothers aspect. Geographics and other aspects such as education are used to predict child care preference. We found that families living in Greater Accra are more likely to send their children to school and the majority of working mothers look after their children by themselves. The analysis of the child care and working mothers data provides insight into current trends and has significant implications for public policy and program decisions."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

This paper uses the Childcare While Working Data [@ghana1998] obtained from the 1998 Ghana Demographic and Health Survey (GDHS). The dataset was collected by downloading the final report of the 1998 GDHS published in the Demographic and Health Surveys (DHS) program. The package 'pdftools' [@citepdftools] is used to download and read the content into R by parsing the pdf. The package 'stringi' [@citestringi] and 'tidyverse' [@citetidyverse] are then used to gather the data.

@@

The questionnaires were made available in the five major local languages (Akan, Ga, Ewe, Hausa, and Dagbani), making the survey accessible to a larger population. After the household listing, spots checks were conducted to make sure the work has been done correctly. Before the actual survey, a test was conducted and adjustments were made to the questionnaires based on the feedback obtained from the test. The questions were comprehensive and detailed to collect information on the socio-economic status of the household. They were also adapted to the laws and norms of the nation.

@@

Since the questionnaires were only available in the five major local languages, if the selected household was from a minor ethnic group and know only their own language, then it would be difficult for the interviewee to understand and answer the survey. Visitors who stayed a night before in the selected household were eligible for the interview. The visitor may be from a different region, so the targeted sample size for each region may be skewed. There is also the possibility of sampling error.

@@

```{r generalchildcare, echo=FALSE, fig.cap="Distribution of the child caretaker for working mothers", fig.width=6, fig.height=4}
totl <-
  ghana %>%
  filter(character=="Total") %>%
  mutate(
    no_child_under_six = round(no_child_under_six/100 * number_of_employed_women, 0),
    one_more_child_under_six = round(one_more_child_under_six/100 * number_of_employed_women, 0)
  )
# make dataset in long format
totl_long <-
  totl %>%
  select(-c(character, total, caretaker_missing)) %>%
  pivot_longer(!background, names_to = "variable", values_to = "count")

@@

# plot the bar graph
totl_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") +
  geom_text(aes(label=paste(percentage, '%'), y=percentage+2),
            size = 3,
            position = position_dodge(width=0.5)) +
  theme_minimal() +
  labs(x='Caretaker', y='Percent') +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other")
  )
```

@@

```{r map, message=FALSE, echo = FALSE, warning=FALSE, fig.cap="Ghana administrative regions map"}
# install.packages('rnaturalearth')
# install.packages('sf')
library(rnaturalearth)
library(sf)
# devtools::install_github("ropensci/rnaturalearthhires")
# The st_centroid function takes the sf object and returns the center of 
# the polygon, again as a sf object.
map_ghana <- ne_states(country = 'Ghana', 
                       returnclass = 'sf')  # grab provinces within Ghana

@@

# plot bar graph
resid_long %>% 
  filter(variables == 'No child under six' | variables == 'One or more child under six') %>% 
  ggplot(aes(x = background, 
             y = percentage,
             fill = variables)) +
  geom_bar(stat = 'identity',
           position = 'dodge') +
  labs(x = "Area",
       y = "Percent",
       fill = "Child status") +
  theme_minimal() +
  scale_fill_brewer() +
  theme(legend.position = "bottom")
```

@@

# plot bar graph
caretaker %>% 
  ggplot(aes(x = caretakers, 
             y = percentage,
             fill = background)) +
  geom_bar(stat = 'identity',
           position = "dodge") +
  labs(x = "Caretaker",
       y = "Percent",
       fill = "Area") +
  theme_minimal() +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

@@

# plot bar graph for caretaker comparisons and administrative regions
child %>% 
  filter(variable != 'number of employed women', variable != 'no child under six', variable != 'one or more child under six') %>% 
  ggplot(aes(x = variable,
             y = percentage,
             fill = background)) +
  geom_bar(stat = 'identity') + 
  labs(x = "Caretaker",
       y = "Percent") +
  theme_minimal() +
  coord_flip() +
  facet_wrap(~background) +
  theme(legend.position = "none")
```

@@

```{r chstatus, echo = FALSE, fig.cap = "Comparing child status in different administrative regions"}
# looking at the number of employed women selected in the different administrative regions
# with the child under six
child %>%
  filter(variable=='no child under six'|variable=='one or more child under six') %>% 
  ggplot(aes(x = background,
             y = percentage,
             fill = variable)) +
  geom_bar(stat = 'identity', position = "dodge", width = 0.8) +
  labs(x = "Administrative regions",
       y = "Percent",
       fill = "Child status") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  coord_flip() 
```

@@

```{r motheredu, echo=FALSE, fig.cap="Comparing the child status of working mothers based on their education level", fig.width=6, fig.height=4}
# select mother edu
mother_edu <-
  ghana %>%
  filter(character=="Mother\'s_education") %>%
  mutate(
    no_child_under_six = round(no_child_under_six/100 * number_of_employed_women, 0),
    one_more_child_under_six = round(one_more_child_under_six/100 * number_of_employed_women, 0)
  )

@@

edu_long %>%
  filter(variable == 'no_child_under_six' | variable == 'one_more_child_under_six') %>%
  ggplot(aes(x = background, y = count, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Mother's education level", y = "Number of employed women", fill = "Child status") +
  theme_minimal() +
  scale_fill_discrete(labels = c("No child under six", 
                                 "One or more child under six")) +
  scale_x_discrete(labels = c("No_education" = "No education")) +
  theme(legend.position = 'bottom')
```

@@

# plot the bar graph
edu_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") + 
  theme_minimal() +
  labs(x='Caretaker', y="Percent") +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired help",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other")
  ) +
  facet_wrap(vars(background),
             dir = "v",
             scales = "free",
             labeller = labeller(background = 
                                   c("No_education" = "No education",
                                     "Primary" = "Primary",
                                     "Middle" = "Middle",
                                     "Secondary+" = "Secondary+")
                                 )
  )
```

@@

```{r workstatus, echo=FALSE, message=FALSE, fig.cap="Comparing the child status of working mothers based on work status", fig.width=6, fig.height=4}
# select work status
work_status <-
  ghana %>%
  filter(character=="Work_status") %>%
  mutate(
    no_child_under_six = round(no_child_under_six/100 * number_of_employed_women, 0),
    one_more_child_under_six = round(one_more_child_under_six/100 * number_of_employed_women, 0)
  )

@@

work_long %>%
  filter(variable == 'no_child_under_six' | variable == 'one_more_child_under_six') %>%
  ggplot(aes(x = background, y = count, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Work status", y = "Number of employed women", fill = "Child status") +
  theme_minimal() +
  scale_x_discrete(labels = c("For_family_member" = "For family member",
                              "For_someone_else" = "For someone else")) + 
  scale_fill_discrete(labels = c("No child under six", 
                                 "One or more child under six")) +
  theme(legend.position='bottom')
```

@@

# plot the bar graph
work_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") + 
  theme_minimal() +
  labs(x='Caretaker', y="Percent") +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired help",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other")
  ) +
  facet_wrap_paginate(vars(background),
                      ncol=2,
                      nrow=2,
                      dir = "h",
                      scales = "free",
                      labeller = labeller(background = c("For_family_member" = "For family member",
                                                         "For_someone_else" = "For someone else",
                                                         "Self-employed" = "Self-employed")
                      )
  )

@@

```{r occupation, echo=FALSE, message=FALSE, fig.cap="Comapring child status based on occupation", fig.width=6, fig.height=4}
# select occupation
occupation <-
  ghana %>%
  filter(character=="Occupation") %>%
  mutate(
    no_child_under_six = round(no_child_under_six/100 * number_of_employed_women, 0),
    one_more_child_under_six = round(one_more_child_under_six/100 * number_of_employed_women, 0)
  )

@@

occup_long %>%
  filter(variable == 'no_child_under_six' | variable == 'one_more_child_under_six') %>%
  ggplot(aes(x = background, 
             y = count,
             fill = variable)) +
  geom_bar(stat = "identity",
           position = "dodge") +
  labs(x = "Sector",
       y = "Number of employed women",
       fill = "Child status") +
  theme_minimal() +
  scale_fill_discrete(labels = c("No child under six", 
                                 "One or more child under six")) +
  theme(legend.position = 'bottom')
```

@@

# plot bar graph
occup_caretaker %>%
  ggplot(aes(x=caretaker, y=percentage, fill=background)) +
  geom_bar(stat = "identity", position="dodge") + 
  labs(x='Caretaker', y="Percent", fill="Occupation") +
  theme_minimal() +
  theme(legend.position='bottom') +
  coord_flip() +
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent", 
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired help",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other")
  )
```

@@

```{r employmentstatus, echo=FALSE, message=FALSE, fig.cap="Comparing child status based on employment status", fig.width=6, fig.height=4}
# select employment status
employment_status <-
  ghana %>%
  filter(character=="Employment_status") %>%
  mutate(
    no_child_under_six = round(no_child_under_six/100 * number_of_employed_women, 0),
    one_more_child_under_six = round(one_more_child_under_six/100 * number_of_employed_women, 0)
  )

@@

emply_long %>%
  filter(variable == 'no_child_under_six' | variable == 'one_more_child_under_six') %>%
  ggplot(aes(x = background, y = count, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Employment status", y = "Number of employed women", fill = "Child status") +
  theme_minimal() +
  scale_fill_discrete(labels = c("No child under six", 
                                 "One or more child under six")) +
  scale_x_discrete(labels = c("All_year_full_week" = "All year, full week",
                              "All_year_part_week" = "All year, part week")) +
  theme(axis.text.x = element_text(angle=15), legend.position = 'bottom')
```

@@

# plot the bar graph
emply_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") + 
  theme_minimal() +
  labs(x='Caretaker', y="Percent") +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired help",
        "child_in_school" = "schooling",
        "other_child" = "siblings",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other")
  ) +
  facet_wrap(vars(background),
             dir = "v",
             scales = "free",
             labeller = labeller(background = c("All_year_full_week" = "All year, full week",
                                                "All_year_part_week" = "All year, part week",
                                                "Seasonal" = "Seasonal",
                                                "Occasional" = "Occasional")
                                 )
  )
